---
title: "Data Analysis 3 : Assignment 1"
author: "Ghazal"
geometry : margin=1.7cm
fontsize: 9pt
output: 
  pdf_document:
  extra_dependencies: ["flafter"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
# Clearing the environment
rm(list=ls())

# Loading the Libraries
library(tidyverse)
library(modelsummary)
library(fixest)
library(kableExtra)
library(data.table)
library(ggplot2)
library(GGally)
library(viridis)
library(caret)
library(grid)
library(dplyr)
library(gridExtra)
library(grid)
library(viridis)

```


```{r message=FALSE, warning=FALSE, include=FALSE}
# Importing the data
data_org <- read_csv("https://osf.io/4ay9x/download")


# Filtering the data for the chosen occupation
dt <- data_org %>% filter(occ2012==630)
dt <-  data.table(dt)

```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Counting the number of observations in each occupation
#aggregate((count = data_org$occ2012), list(value = data_org$occ2012), length)


```


```{r message=FALSE, warning=FALSE, include=FALSE}



# Creating new variable for earnings per hour and log of it
dt <- dt %>% 
  
  # Create : Female dummy variable
  mutate(female=as.numeric(sex==2)) %>% 
  
  # Create : hourly wage (earning per number of hours worked per week)
  mutate(w = earnwke/uhours) %>%
  
  # Add : log of wage
  mutate(lnw=log(w)) %>% 
  
  # Filter : Education is School's degree or higher
  filter(grade92 >= 38) %>% 
  
  # Filter : Age is above 20 years old
  filter(age>=18) %>% 
  
  # Filter : wage more than USD 1
  filter(w>1) %>% 
  
  # Create : Education Dummies
  mutate(
  no_degree=as.numeric(grade92<=40),
  ed_associate=as.numeric(grade92==41 |grade92==42) ,
  ed_BA=as.numeric(grade92==43),
  ed_MA=as.numeric(grade92==44),
  ed_Profess = as.numeric(grade92==45),
  ed_PhD = as.numeric(grade92==46))

dt <- dt[grade92 <= 40, educ := "No Degree"][grade92==41 | grade92==42, educ := "Associate" ][grade92 == 43, educ := "Bachelors"][grade92 == 44, educ := "Masters"][grade92 == 45, educ := "Professional"][grade92 == 46, educ := "PhD"]

# Looking for education data summary
sum1 <- datasummary(w * (`Education` = as.factor(educ)) ~ N + Percent() + mean, data = dt, title = "Wage and education summary")



# Looking for calss data summary
sum2 <- datasummary(w* (`class` = as.factor(class)) ~ N + Percent() + mean, data = dt, title = "Wage and class summary")  


    # Create age squared
dt <-  dt %>% mutate(agesq=age^2) %>% 
  
  # Create Private Sector Dummy
  mutate(privt_sec = as.numeric(class=="Private, Nonprofit" | class== "Private, For Profit"))



# Looking for male and female data summary
datasummary(w*factor(sex) ~ N + Percent() + mean, data = dt)


  # Create if individuals own child or no

dt <- dt %>% 
  mutate(child = as.numeric(ownchild >= 1))

# Looking for hourly wage of individuals with children and without children data summary
sum3 <- datasummary(w * (`Own a child` = factor(child)) ~ N +Percent() + mean, data = dt, title = "Wage and Own a child summary")


dt <- dt[sex == 1, gender := "Male"][sex == 2, gender := "Female"]

# Looking for male and female data summary
sum4 <- datasummary(w*(`Gender` = as.factor(gender)) ~ N + Percent() + mean, data = dt, title = "Wage and Gender summary")

# Looking for data summary of individuals per each gender with children
sum5 <- datasummary(w * (`Gender` = as.factor(gender)) * (`Own a child` = factor(child)) ~ N + Percent() + mean, data = dt, title = "Wage, Gender and own a child summary")

# Checking hourly wage education level and gender data summary
sum6 <- datasummary(w* (`Education` = as.factor(educ))* (`Gender` = as.factor(gender)) ~ N + Percent() + mean, data = dt, title = "Wage, education and Gender summary")

# Checking race summary
datasummary(as.factor(race) ~ N, data = dt)

# Checking race summary
datasummary(as.factor(marital) ~ N, data = dt)

# Add new variable for marital status, if married then 1 else 0
dt$marital_status <- ifelse(dt$marital <= 3, 1, 0)

sum7 <- datasummary(w* (`Married` = factor(marital_status)) ~ N + Percent() + mean, data = dt, title = "Wage and Marital status summary")
sum8 <- datasummary(w* (`Education` = as.factor(educ))*(`Married` = as.factor(marital_status)) ~ N + Percent() + mean, data = dt, title = "Wage, education and marital status summary")

sum9 <- datasummary(w* (`Gender` = as.factor(gender))*(`Married` = as.factor(marital_status)) ~ N + Percent() + mean, data = dt, title = "Wage, gender and marital status summary")
```




```{r include=FALSE}

# Figures are put in the Appendix


# Earnings per hour histogram
hist_w <- ggplot(data=dt, aes(x=w)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)), binwidth = 5, boundary=0, fill = "#440154", color = "white", size = 0.2, alpha = 0.5,  show.legend=F, na.rm=TRUE) +
  labs(x = "Earnings per hour",y = "Percent")+
  theme_bw() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_y_continuous(expand = c(0.01,0.01),labels = scales::percent_format(accuracy = 1)) +
  ggtitle("Earnings per hour") +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))

# Earnings per hour ln histogram
hist_w_ln <- ggplot(data=dt, aes(x=lnw)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)), boundary=0, fill = "#440154", color = "white", size = 0.2, alpha = 0.5,  show.legend=F, na.rm=TRUE) +
  labs(x = "Earnings per hour Ln (ln(w))",y = "Percent")+
  theme_bw() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_y_continuous(expand = c(0.01,0.01),labels = scales::percent_format(accuracy = 1)) +
  ggtitle("Earnings per hour Ln ") +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))


# Male hourly wage histogram
hist_w_m <- ggplot(data=dt[sex == 1], aes(x=w)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)), binwidth = 5, boundary=0, fill = "#440154", color = "white", size = 0.2, alpha = 0.5,  show.legend=F, na.rm=TRUE) +
  labs(x = "Male Earnings per hour",y = "Percent")+
  theme_bw() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_y_continuous(expand = c(0.01,0.01),labels = scales::percent_format(accuracy = 1)) +
  ggtitle("Male Earnings per hour ") +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))

# Female hourly wage histogram
hist_w_f <- ggplot(data=dt[sex == 2], aes(x=w)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)), binwidth = 5, boundary=0, fill = "#440154", color = "white", size = 0.2, alpha = 0.5,  show.legend=F, na.rm=TRUE) +
  labs(x = "Female Earnings per hour",y = "Percent")+
  theme_bw() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_y_continuous(expand = c(0.01,0.01),labels = scales::percent_format(accuracy = 1)) +
  ggtitle("Female Earnings per hour") +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))

# Male age histogram

hist_age_m <- ggplot(data=dt[sex == 1], aes(x=age)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)), binwidth = 5, boundary=0, fill = "#440154", color = "white", size = 0.2, alpha = 0.5,  show.legend=F, na.rm=TRUE) +
  labs(x = "Male Age Distribution",y = "Percent") +
  theme_bw() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_x_continuous(breaks = seq(15,70, 5)) +
  scale_y_continuous(expand = c(0.01,0.01),labels = scales::percent_format(accuracy = 1)) +
  ggtitle("Male age distribution") +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))

# Female age historam
hist_age_f <- ggplot(data=dt[sex == 2], aes(x=age)) +
  geom_histogram(aes(y = (..count..)/sum(..count..)), binwidth = 5, boundary=0, fill = "#440154", color = "white", size = 0.2, alpha = 0.5,  show.legend=F, na.rm=TRUE) +
  labs(x = "Female Age Distribution",y = "Percent") +
  theme_bw() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_x_continuous(breaks = seq(15,70, 5)) +
  scale_y_continuous(expand = c(0.01,0.01),labels = scales::percent_format(accuracy = 1)) +
  ggtitle("Female age distribution") +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))

# Box plots and jitter
# Box plot of hourly wage for each gender
jitter_gender <- ggplot(dt,aes(factor(gender), w, color = "#440154" )) + 
  geom_boxplot(alpha = 0.1, frame = FALSE) + 
  geom_jitter(height = 0, width = 0.1, alpha = 0.2) +
  scale_color_viridis(option = "D", discrete = TRUE)+
  scale_fill_viridis(option = "D", discrete = TRUE) +
  theme_bw() +
  theme(legend.position = "none", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5) ) +
  labs(x = "Gender", y = "Earnings per hour") +
  ggtitle("Gender Earning per hour ")

level_educ <- factor(dt$educ, level = c("No Degree", "Associate","Bachelors", "Masters", "Professional", "PhD"))


# Box plot of hourly wage for levels of education
jitter_edu <- ggplot(dt,aes(factor(level_educ), w, color = "#440154" )) + 
  geom_boxplot(alpha = 0.1, frame = FALSE) + 
  geom_jitter(height = 0, width = 0.1, alpha = 0.2) +
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  theme_bw() +
  theme(legend.position = "none", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  labs(x = "Education", y = "Earning per hour") + 
  ggtitle("Education Earning per hour")
  #scale_x_discrete(labels = c("Bachelors", "Masters", "Professional", "PhD"))

# Box plot of hourly wage for work sector
jitter_class <- ggplot(dt,aes(factor(class), w, color = "#440154" )) + 
  geom_boxplot(alpha = 0.1, frame = FALSE) + 
  geom_jitter(height = 0, width = 0.1, alpha = 0.2) +
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  theme_bw()+
  theme(legend.position = "none", axis.text.x = element_text(angle= 45, vjust=.5), panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  labs(x = "Working Sector", y = "Earning per hour") +
  ggtitle("Sectors Earning per hour")

# Box plot for hourly wage if the individual has a child or no
jitter_child <- ggplot(dt,aes(factor(child), w, color = "#440154" )) + 
  geom_boxplot(alpha = 0.1, frame = FALSE) + 
  geom_jitter(height = 0, width = 0.1, alpha = 0.2) +
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  theme_bw() +
  theme(legend.position = "none", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  scale_x_discrete(labels = c("No", "Yes")) +
  labs(x = "Child", y = "Earnings per hour") +
  ggtitle("Own child Earning per hour")

# Box plot for hourly wage if the individual is married or not
jitter_married <- ggplot(dt,aes(factor(marital_status), w, color = "#440154" )) + 
  geom_boxplot(alpha = 0.1, frame = FALSE) + 
  geom_jitter(height = 0, width = 0.1, alpha = 0.2) +
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  theme_bw() +
  theme(legend.position = "none", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  scale_x_discrete(labels = c("No", "Yes")) +
  labs(x = "Married", y = "Earning per hour") +
  ggtitle("Marital Status Earning per hour")

# Two variables boxplots
# Box plot of hourly wage for education level per gender 
boxplot_edu_gender <- ggplot(dt, aes(x = factor(level_educ), y = w, fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.6, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  labs(x = "Education Levels",y = "Earning per hour in USD", color = "Gender", fill = "Gender") + 
  theme_bw() +
  theme(legend.position = "top", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5) ) +
  ggtitle("Gender and education levels Earning per hour ")

# Box plot of hourly wage for education level per marital status 
boxplot_edu_married <- ggplot(dt, aes(x = factor(level_educ), y = w, fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.6, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  labs(x = "Education Levels",y = "Earning per hour in USD", color = "Gender", fill = "Gender") + 
  theme_bw() +
  theme(legend.position = "top", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  ggtitle("Marital status and education level Earning per hour")
  

# Box plot of hourly wage for child per gender 
boxplot_child_gender <- ggplot(dt, aes(x = factor(child), y = w, fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.6, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  labs(x = "Child",y = "Earning per Hour in USD", color = "Gender", fill = "Gender") + 
  theme_bw() +
  theme(legend.position = "top", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5) ) +
  scale_x_discrete(labels = c("No", "Yes")) +
  ggtitle("Gender and owning a child Earning per hour")

# Box plot of hourly wage for Working sector per gender
boxplot_class_gender <- ggplot(dt, aes(x = factor(class), y = w, fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.6, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  labs(x = "Working Sector",y = "Earning per Hour in USD", color = "Gender", fill = "Gender") + 
  theme_bw() +
  theme(legend.position = "top", axis.text.x = element_text(angle= 45, vjust=.5), panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))+
  ggtitle("Gender and sectors Earning per hour")

# Box plot of hourly wage for marital status per gender
boxplot_married_gender <- ggplot(dt, aes(x = factor(marital_status), y = w, fill = factor(gender), color=factor(gender))) +
  geom_boxplot(alpha=0.6, na.rm=T, outlier.shape = NA, width = 0.8) +
  stat_boxplot(geom = "errorbar", width = 0.8, size = 0.3, na.rm=T)+
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  labs(x = "Married",y = "Earning per Hour in USD", color = "Gender", fill = "Gender") + 
  theme_bw() +
  theme(legend.position = "top", panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  scale_x_discrete(labels = c("No", "Yes")) +
  ggtitle("Gender and marital status Earning per hour")


# loess plot
# loess for age and Earning per hour

scatter_age <- ggplot( dt , aes(x = age, y = w)) +
  geom_point(size=0.5,alpha=0.6, na.rm = T, color = "#440154") +
  geom_smooth(method="loess" , formula = y ~ x , na.rm = T, color = "#fde725" )+
  theme_bw() +
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  labs(x = "Age",y = "Wage per Hour in USD", color = "Gender", fill = "Gender") +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  ggtitle("Loess with age")
# 

scatter_edu <- ggplot( dt[grade92 != 38] , aes(x = grade92, y = w)) +
  geom_point(size=0.9, alpha=0.8, na.rm = T, color = "#440154") +
  geom_smooth(method="loess" , formula = y ~ x , na.rm = T, color = "#fde725")+
  theme_bw() +
  scale_color_viridis(option = "D", discrete = TRUE) +
  scale_fill_viridis(option = "D", discrete = TRUE) +
  labs(x = "Education Levels",y = "Wage per Hour in USD") +
  #scale_x_continuous(labels = c("High school", "College", "Associate Vocational", "Associate Academic", "Bachelors", "Masters", "Professional", "PhD")) +
  theme(panel.border = element_blank(), axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5)) +
  ggtitle("Loess with education")


```


### 1.Introduction

The purpose of this assignment is to build four predictive models using linear regression. The source of data for this purpose is CPS-earning data set which can be found [here](https://osf.io/g8p9j/). The chosen occupation for this project is _Human Resource worker_ with the census occupation code of _630_ which has 701 observations. GitHub link can be found [here](https://github.com/ghazalayobi/DA3).

### 2. Data Munging and Transformation 

Predicting the weekly earning is a quantitative prediction exercise. For the process of data munging and transformation, hourly wage_**(w)**_ is calculated by dividing the weekly earnings (earnwke) by the number of hours (uhours) and the log of the mentioned variable _**(lnw)**_ is also created. To model other variables are created and transformed which are as following: education level, marital status, if an individual owns a child, working sector dummies are created. Other character variables such as Gender and education levels are created. To model non-linearity in age for the regression with earning per week as a dependent variable, quadratic age predictor is created. The sample is varied with high number of females (521) compared to men (180). 
For further data filtering process, education is included from high school to PhD. The education levels below college are added into _Non-degree_ category, and both associate certificates of vocational and academic are added to _associate_. The data contains observation who are 18 years old or above and less than 64 years. Moreover, the selected data set is transformed to observations who earns more than one USD per week. 

### 3. Variables, Interations, and Regressions

**Variables :** Education is likely a strong predictor of earning per hour. Table 4, in the Appendix indicates that the mean difference between of a human resource worker with no degree to a human resource worker who holds a PhD degree is 27 USD. Moreover, other variables such as age and age square are added to the predication models. Gender plays an important role, Table 7 shows that female human resource workers earn 4 USD less than their male counterparts. Moreover, a binary variable capturing if human resource worker is married or otherwise, indicates interesting summary in the Appendix Table 10. It shows that married human resource workers earn more than the otherwise. 
**Interactions :** To further capture the interplay of independent variables, interactions are used. Interactions such as gender and education, marital status and gender, gender times owning a child and education are used to understand the interplay of variables. **Regressions :** Four linear regression models are built to prediction analysis. As Table 1, in the appendix shows, Model 1 is the simplest containing education dummies, as following Model 2 has the Model 1 explanatory variables along with age and age squared. Moreover, in the Model 3, more explanatory variables such as gender, working sector, marital status, and own-child are added. Model 4, is the most complex among all model. This mentioned model contains all the mentioned independent variables and the respective interactions.  

### 4. Model Performance

BIC is the measure of the fit of a model using all the original data and it penalizes the model complexity and helps to avoid over-fitting. Models with lower BIC are generally preferred. Among the models Model 2 has the lowest BIC, however it has minimal difference with Model 3. The second measurement to evaluate model performance is RMSE, which is the average squared loss across several target observations. RMSE is the lowest for the models 3 and 4. By looking to the Table 2 for the result of cross validation of the models, it suggests that Model 2 and Model 3 have the best properties. Model 3 has the lowest for BIC and cross validation RMSE average. Thus, Model 3 is selected for the purpose of the project which contains 14 variables.


```{r include=FALSE}

# Creating 5% and 95% 
P05 <- function(x){ quantile(x,.05,na.rm=T)}
P95 <- function(x){ quantile(x,.95,na.rm=T)}

# Data Summary
summary_table <- datasummary(
  (`Weekly earnings` = earnwke) + 
  (`Weekly hours worked` = uhours) + 
  (`Earning per hour` = w) + 
  (`Female` = female) +
  (`No degree` = no_degree) +  
  (`Associate` = ed_associate) +
  (`BA Degree` = ed_BA) + 
  (`MA Degree` = ed_MA) + 
  (`Professional Degree` = ed_Profess) + 
  (`PhD` = ed_PhD) + 
  (`Age` = age) + 
  (`Work in Private Sector` = privt_sec) + 
  (`Has child` = child) +
  (`Marital Status` = marital_status) ~ (Median + Mean + SD + Min + Max + P05 + P95 + N), 
  data = dt, 
  title = "Descriptive Summary Statistics" ) %>% 
  kable_styling(latex_options = c("HOLD_position","scale_down"))


```


```{r include=FALSE}
# Regressions
# Simple Model one

m1 <- as.formula(w ~ ed_associate + ed_BA + ed_MA + ed_Profess + ed_PhD)
m2 <- as.formula(w ~ ed_associate + ed_BA + ed_MA + ed_Profess + ed_PhD + age + agesq)

m3 <- as.formula(w ~ ed_associate + ed_BA + ed_MA + ed_Profess + ed_PhD + age + agesq + gender + class + child + marital_status)

m4 <- as.formula(w ~ ed_associate + ed_BA + ed_MA + ed_Profess + ed_PhD + age + agesq + female + class + marital_status + ed_associate *female  + female * ed_BA + female * ed_MA + female * child * educ)

reg1 <- feols(m1, data = dt, vcov = 'hetero')
reg2 <- feols(m2, data = dt, vcov = 'hetero')
reg3 <- feols(m3, data = dt, vcov = 'hetero')

# Complex Regression
reg4 <- feols(m4, data = dt, vcov = 'hetero')

```

```{r include=FALSE}

# Adding variables names
variable_names <- c('(Intercept)' = 'Intercept',
                    'w' = 'Hourly wage',
                    'ed_associate' = 'Associate',
                    'ed_BA' = 'BA Degree',
                    'ed_MA' = 'MA Degree',
                    'ed_Profess' = 'Professional Degree',
                    'ed_PhD' = 'PhD',
                    'age' = 'Age',
                    'agesq' = 'Age Squared',
                    'female' = 'Female',
                    'privt_sec' = 'Private Sector',
                    'child' = 'Has Child',
                    'marital_status' = 'Married')
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# evaluation of the models: using all the sample


fitstat_register("k", function(x){length( x$coefficients ) - 1}, "No. Variables")
etable( 'M1' = reg1 , 'M2' = reg2 , 'M3' = reg3 , 'M4' = reg4, dict = variable_names , fitstat = c('aic','bic','rmse','r2','n','k') ) %>% kbl(caption = "Regression Models for Earning per hour", booktabs = T) %>% kable_styling(full_width = T)
```

```{r include=FALSE}

# Simple k-fold cross validation setup:
# set number of folds to use (must be less than the no. observations)
k <- 4

# Model 1
set.seed(1238)
cv1 <- train(m1, dt, method = "lm", trControl = trainControl(method = "cv", number = k))

# Model 2
set.seed(1238)
cv2 <- train(m2, dt, method = "lm", trControl = trainControl(method = "cv", number = k))

# Model 3
set.seed(1238)
cv3 <- train(m3, dt, method = "lm", trControl = trainControl(method = "cv", number = k))

# Model 4
set.seed(1238)
cv4 <- train(m4, dt, method = "lm", trControl = trainControl(method = "cv", number = k))


# Calculate RMSE for each fold and the average RMSE as well
cv <- c("cv1", "cv2", "cv3", "cv4")
rmse_cv <- c()

for(i in 1:length(cv)){
  rmse_cv[i] <- sqrt((get(cv[i])$resample[[1]][1]^2 +
                       get(cv[i])$resample[[1]][2]^2 +
                       get(cv[i])$resample[[1]][3]^2 +
                       get(cv[i])$resample[[1]][4]^2)/4)
}


# summarize results
cv_mat <- data.frame(rbind(cv1$resample[4], "Average"),
           rbind(cv1$resample[1], rmse_cv[1]),
           rbind(cv2$resample[1], rmse_cv[2]),
           rbind(cv3$resample[1], rmse_cv[3]),
           rbind(cv4$resample[1], rmse_cv[4])
           )

colnames(cv_mat)<-c("Resample","Model1", "Model2", "Model3", "Model4")
cv_mat <- cv_mat %>% kbl(caption = "4-Fold Cross Validation and RMSE", booktabs = T) %>% kable_styling(full_width = T)

```



```{r echo=FALSE}
# Model complexity and out-of-sample RMSE performance
m_comp <- c()
models <- c("reg1", "reg2", "reg3", "reg4")
for( i in 1 : length(cv) ){
  m_comp[ i ] <- length( get( models[i] )$coefficient  - 1 ) 
}

m_comp <- tibble( model = models , 
                  complexity = m_comp,
                  RMSE = rmse_cv )

model_performance <- ggplot( m_comp , aes( x = complexity , y = RMSE ) ) +
  geom_point(color='red',size=2) +
  geom_line(color= "#440154",size=0.5)+
  labs(x='Number of explanatory variables',y='Averaged RMSE on test samples',
       title='Prediction performance and model compexity') +
  theme_bw() +
  theme(axis.text=element_text(size=8), plot.title = element_text(size = 12L, face = "bold", hjust = 0.5))


```



\newpage
# Appendix



```{r echo=FALSE, message=FALSE, warning=FALSE, show.fig = "hold", fig.height=4, fig.width=8}

# Descriptive Data Summary

cv_mat
grid.arrange( model_performance, ncol = 1, bottom = "Cross-Validation RMSE in the graphs shows lowest result for the Model 3. This Model contains 14 Variables")


```

\newpage
# Descriptive Data Summary

```{r echo=FALSE}
summary_table


```

Descriptive summary of the main variables in the data set can be in the above table. From the table it can be inferred that because of the presence of high hourly wage values like USD 2884, mean tends to be to the right of the median thus making the sample distribution rightly skewed. Similarly, we see that there are certain people who work more than 40 hours (maximum value of 65 hours a week) which is also the cause for skewness. Moreover, there is also the presence of extreme values. For example, the minimum wage value is computed out to be USD 2.31

\newpage
# Histograms

```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, fig.height=9, fig.width=8}

# Histograms

grid.arrange( grobs = list(hist_w, hist_w_ln, hist_w_f, hist_w_m, hist_age_f, hist_age_m), ncol = 2, bottom = )

```


\newpage
# One variable plots


```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, fig.height=9, fig.width=8}

grid.arrange( grobs = list(jitter_gender, jitter_child, jitter_edu, jitter_married), ncol = 2)

```

\newpage
# One variable plot

```{r echo=FALSE, message=FALSE, warning=FALSE}
jitter_class

```



\newpage
# Two Variable plots

```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, fig.height=9, fig.width=8}

grid.arrange( grobs = list(boxplot_edu_gender, boxplot_married_gender, boxplot_edu_married, boxplot_child_gender), ncol = 2)

```

\newpage
# Two Variable plot

```{r echo=FALSE, message=FALSE, warning=FALSE}
boxplot_class_gender
```


\newpage
# Loess

```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, show.fig = "hold"}

grid.arrange( grobs = list(scatter_edu, scatter_age), ncol = 2)

```

\newpage


```{r echo=FALSE, message=FALSE, warning=FALSE, show.fig = "hold", fig.align = "center"}


sum1
sum2
sum3
sum4
sum5
sum6
sum7
sum8
sum9
```









